---
title: "Stability EEG Ketamine Dose Classifier"
output: html_notebook
author: Diego G. Davila, PhD Candidate, Proekt Lab, University of Pennsylvania School of Medicine
---

This notebook covers the development of a classifier trained to pick out a subject's ketamine blood concentration based on linear stability measures derived from scalp EEG. 

# Setup
```{r}
library(pacman)
p_load(dplyr, tidyverse, glmnet, randomForest, tree, ggplot2, skimr, ggRandomForests, xgboost, kableExtra, keras, neuralnet, tensorflow, ggpubr, effectsize, pROC, verification, MASS, pracma, ROCR, e1071)
```

# Smoothed Criticality Index Histogram Model

## Load data
```{r}
data2 <- read.csv('EEG_CSV/classification_lambda_histograms.csv', header = FALSE)
data2[,195] <- as.factor(data2[,195]) # Turn the response variable into a factor
data2[,196] <- as.factor(data2[,196]) # subject id
colnames(data2)[196] <- c("subject.id")

#skimr::skim(data2)
```


## Create and train the classifier
```{r}
rf_classifier_2 <- randomForest(data = data2,        # Pass in the data
                              x = data2[,151:194],   # Select the predictor variables (0.8-1.2 Crit. Index)
                              y = data2[,195],       # Select the response variable
                              ntree = 500,           # Select the range of number of trees to try
                              mtry = 6)             # Select the number of variables for each tree to use
```

## Plotting
```{r}
oob.err_2 <- gg_error(rf_classifier_2)
colnames(oob.err_2) <- c('All','0.0μg/ml', '0.4μg/ml', 'ntree')

plot(oob.err_2) + 
  ggtitle("Stability-Based RandomForest Classifier: Ketamine Dose") 


# Plot the feature importance
jpeg(file="C:/Users/diego/Desktop/EEGStabilityGroupFigures/Classifier_FeatureImportance.jpeg", width = 300, height = 200, units='mm', res = 300)
rf_2_feature_importance <- data.frame(importance(rf_classifier_2), Criticality_Bin = seq(0.8, 1.015, by = 0.005))
ggplot(data = rf_2_feature_importance, aes(x=Criticality_Bin, y=MeanDecreaseGini)) + 
  geom_bar(stat="identity", fill="darkcyan") +
  ggtitle("Stability Classifier Feature Importance") + 
  theme(plot.title = element_text(face="bold", size=15), axis.title = element_text(face="bold", size=15), axis.text = element_text(face="bold", size=15)) 
dev.off()
```

## Surrogate Data Testing

### Load Data
```{r}
surrogate <- read.csv('EEG_CSV/surrogate_classification_lambda_histograms.csv', header = FALSE)
surrogate[,195] <- as.factor(surrogate[,195]) # Turn the response variable into a factor
surrogate[,196] <- as.factor(surrogate[,196]) # subject id
colnames(surrogate)[196] <- c("subject.id")
#skimr::skim(surrogate)
```

### Create and train the classifier
```{r}
rf_classifier_4 <- randomForest(data = surrogate,        # Pass in the data
                              x = surrogate[,151:194],   # Select the predictor variables (0.8-1.2 Crit. Index)
                              y = surrogate[,195],       # Select the response variable
                              ntree = 500,           # Select the range of number of trees to try
                              mtry = 6)             # Select the number of variables for each tree to use 
```

### Plotting
```{r}
oob.err_4 <- gg_error(rf_classifier_4)
colnames(oob.err_4) <- c('All','0.0μg/ml', '0.4μg/ml', 'ntree')

plot(oob.err_4) + 
  ggtitle("Stability-Based RandomForest Classifier: Ketamine Dose (Surrogate)") 

rf_4_feature_importance <- data.frame(importance(rf_classifier_4), Criticality_Bin = seq(0.8, 1.015, by = 0.005))

ggplot(data = rf_4_feature_importance, aes(x=Criticality_Bin, y=MeanDecreaseGini)) + 
  geom_bar(stat="identity", fill="darkcyan") +
  ggtitle("Classifier Feature Importance: Surrogate Data")
  
```


# Leave-one-out testing for each subject

## Original Data
```{r}
# store prediction errors here
prediction.error <- rep(0,6)

for (i in 1:6){
  train.x <- data2[(as.numeric(data2$subject.id) != i),]
  test.x <- data2[(as.numeric(data2$subject.id) == i),]
  
  # train the model
  rf_classifier_x <- randomForest(data = train.x,        # Pass in the data
                              x = train.x[,151:194],   # Select the predictor variables (0.8-1.2 Crit. Index)
                              y = train.x[,195],       # Select the response variable
                              ntree = 500,           # Select the range of number of trees to try
                              mtry = 6)             # Select the number of variables for each tree to use 
  
  pred.x <- predict(object = rf_classifier_x, newdata = test.x[,151:194]) # Get prediction for left-out subject
  prediction.error[i] <- sum(pred.x != test.x[,195])/dim(test.x)[1] # calculate the error rate
} 
prediction.error
```

```{r}
# Plot Leave-One-Out Error By Subject
pe <- data.frame(prediction.error)
pe$Left_Out_Subject <- c("Subject 1", "Subject 2", "Subject 3", "Subject 4", "Subject 5", "Subject 6")
colnames(pe)[1] <- c("Prediction_Error")
jpeg(file="C:/Users/diego/Desktop/EEGStabilityGroupFigures/Stability_LeaveOneOut_PredictionError.jpeg", width = 300, height = 200, units='mm', res = 300)
ggplot(data = pe, aes(y=Prediction_Error, x=Left_Out_Subject))+
  geom_col(fill='aquamarine4') + 
  ylim(c(0, 1)) +
  ggtitle("Stability Classifier 'Leave-One-Out' Prediction Error") +
  xlab("Left-out Subject") +
  ylab("Prediction Error") + 
  geom_text(aes(label=sprintf("%0.3f", round(Prediction_Error, digits = 3))), position=position_dodge(width=0.9), vjust=-0.25) +
  theme(plot.title = element_text(face="bold", size=15), axis.title = element_text(face="bold", size=15), axis.text = element_text(face="bold", size=15))
dev.off
```


## Surrogate Data
```{r}
# store prediction errors here
prediction.error.surr <- rep(0,6)

for (i in 1:6){
  train.x <- surrogate[(as.numeric(surrogate$subject.id) != i),]
  test.x <- surrogate[(as.numeric(surrogate$subject.id) == i),]
  
  # train the model
  rf_classifier_x <- randomForest(data = train.x,        # Pass in the data
                              x = train.x[,151:194],   # Select the predictor variables (0.8-1.2 Crit. Index)
                              y = train.x[,195],       # Select the response variable
                              ntree = 500,           # Select the range of number of trees to try
                              mtry = 6)             # Select the number of variables for each tree to use 
  
  pred.x.surr <- predict(object = rf_classifier_x, newdata = test.x[,151:194]) # Get prediction for left-out subject
  prediction.error.surr[i] <- sum(pred.x.surr != test.x[,195])/dim(test.x)[1] # calculate the error rate
} 
prediction.error.surr
```

## Prediction Error Table
```{r}

all.errors <- data.frame(Real = prediction.error, Surrogate = prediction.error.surr)
rownames(all.errors) <- c("Subject 1", "Subject 2", "Subject 3", "Subject 4", "Subject 5", "Subject 6")
kable(all.errors, align='cc', format = 'html')

```

